Learning and calibrating per-location classifiers for visual place recognition\thanks{

### 2015-12-04 manuscript.tex
references: consistency - shortstrings references: dual citations introduction: ?? issue introduction: location and place issue paper: terminology consistency

\author{Petr Gron{\'a}t         \and
        Josef {S}ivic          \and
        Tom{\'a}{\v s} Pajdla   \and
        Guillaume Obozinski%etc.
}

@@

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}
\vspace{-5mm}
\date{Received: date / Accepted: date}

@@

  In this work, we also take advantage of geotags as an available form of supervision and investigate whether the place recognition problem can be cast as a classification task.
  {
    Learning visual classifiers has been investigated for landmark recognition~\cite{Li09} where consumer photographs were clustered into landmark classes based on geo-tags. 
    In this work we wish to recognize individual street locations rather than a small number of landmarks, and as a consequence have only a few (1-5) photographs
    %available for each place.   
    \textcolor{petr}{capturing the same location}.
    In particular, we train a classifier {\em for each location on the map} in a similar manner to per-exemplar classification in object recognition~\cite{Malisiewicz11}. 
    \textcolor{petr}{
      By \em{location} we mean a scene captured by a camera, hence using our terminology, a camera heading to the north captures different location than a camera heading to the south, even though these two cameras share the same GPS location. 
    }
  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Per-location classifiers for place recognition}
\label{sec:classifiers}
%%%%%%%%%%%%%%%%%%%%%%%%
   We are given an image descriptor $\vec{x}_j$, one for each database image $j$. This representation can be a sparse tf-idf weighted bag-of-visual-words vector~\cite{Sivic03} or a dense compact descriptor such as the Fisher vector (FV)~\cite{Jegou12}. The goal is to learn 
   a score \textcolor{petr}{function}
   $f_j$ for each database image $j$, so that, at test time, given the descriptor $\vec{q}$ of the query image, we can either retrieve the correct target image as the image $j^*$ with the highest score
   % 
   \begin{equation}
   \label{eq:class}
    j^*=\operatorname*{arg\;max}_{j} f_j(\vec{q}) 
   \end{equation}
    %
   \noindent
   or use these scores to rank candidate images and use geometric verification to identify the correct location in an $n$-best list.
   Instead of approaching the problem directly as a large multiclass classification problem, we tackle the problem by learning a per-exemplar linear SVM classifier~\cite{Malisiewicz11}  for each database image $j$.
   Similar to~\cite{Knopp2010}, we use the available geotags to construct the negative set $\mathcal N_j$ for each image $j$. The negative set is constructed so as to concentrate difficult negative examples, i.e.\ from images that are far away from the location of image $j$ and similar to the target image as measured by the dot product between their feature vectors. The details of the construction procedure will be given in section~\ref{sec:experiments}.  The positive set $\mathcal P_j$ is represented by a single positive example
   , which is $\vec{x}_j$ itself. 
   \textcolor{petr}{
    For some locations the positive set can be expanded, details of this procedure will be given in~\ref{sub:parametersOfPerLoc}.
   }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%  
\section{Non-parametric calibration of the  SVM-scores from negative examples only}
\label{sec:calibration}
%%%%%%%%%%%%%%%%%%%%%%%%%%  
  In this section we describe a classifier calibration method that exploits the availability of large amounts of negative data, i.e.\ images from other far away locations in the database.
  In particular, the method estimates the significance of the score of a test example compared to the typical score of  the (plentifully available) negative examples. Intuitively, we will use a large dataset of negative examples to calibrate the individual classifiers so that they {\em reject the same number of negative examples} at each level of the calibrated score.   We will expand this idea in detail using the concepts from hypothesis testing.

   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \subsection{Calibration \emph{via} significance levels}
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      In the following, we view  the problem of deciding whether a query image matches a given location based on the corresponding SVM score as a hypothesis testing problem. In particular, we appeal to ideas from the traditional frequentist hypothesis testing framework also known as Neyman-Pearson (NP) framework (see e.g.~\cite{casella2001statistical}, chap.~8).

@@

      In the NP framework, the \emph{significance level} of a score is measured by the p-value or equivalently by the value of the cumulative density function  (cdf) of the distribution of the negatives at a given score value. The cdf is the function $F_0$ defined by $F_0(s)=\PP(S_0\leq s)$, where $S_0$ is a random variable corresponding to the scores of negative data (see figure~\ref{fig:qntExample} for an illustration of the relation between the cdf and the density of the function). The cdf 

@@

      where $(s_n)_{1\leq n \leq \Ncal }$ are the SVM scores associated with $\Ncal$ negative examples used for calibration. Note that no positive examples are involved in the construction of the cumulative density function.

@@

      We also interpolate the empirical cdf between consecutive datapoints so that instead of being a staircase function it is a continuous piecewise linear function such as illustrated in figure~\ref{fig:calib}. Given a query, we first compute its SVM score $s_q$ and then compute the calibrated probability $f(q)=\hat{F}_0(s_q)$.
      We obtain a similar calibrated probability $f_j(q)$ for each of the SVMs associated with each of the target locations, which can now be ranked.

@@

    For each trained place-specific classifier $s_j$ we construct the empirical cumulative density function~\eqref{eq:empiricalCDF} of scores of the negative examples and keep only its top $K$ values. This can be done offline and the procedure is summarized in Algorithm \ref{alg:offline}. 

@@

   %%%%%%%%%%%%%%%%%%%%%%%
   \subsection{Discussion}
   \label{sec:discussion}
   %%%%%%%%%%%%%%%%%%%%%%% 
      It should be noted that basing the calibration only on the negative data has the advantage that we privilege precision over recall, which is justified given the imbalance of the available training data (many more negatives than positives).
      Indeed, since we are learning with a single positive example, intuitively, we cannot guarantee that the learned partition of the space will generalize well to other positives, whose scores in the test set can potentially drop significantly. % (this is indeed what we observe in practice). 
      By contrast, since we are learning from a comparatively large number of negative examples, we can trust the fact that new negative examples will stay in the half-space containing the negative training set, so that their scores are very unlikely to be large. Our method is therefore based on the fact that we can measure reliably how surprising a high score would be if it was the score of a negative example. This exactly means that we can control false positives (type I error) reasonably well but not false negatives (type II error or equivalently the power of our test/classifier), exactly as in the Neyman-Pearson framework.

      An additional reason for not relying on positive examples for the calibration in our case is that (even if we had sufficiently many of them) the positive examples that we collect using location and geometric verification from the geotagged database typically have illumination conditions that are extremely similar to each other and not representative of the distribution of test positives which can have very different illuminations. This is because of the controlled nature of the capturing process of geotagged street-level imagery (e.g.\ Google Street View) used for experiments in this work. Close-by images are typically captured at a similar time (e.g. on the same day) and under similar imaging conditions.

@@

  \subsection{Affine calibration model}
    Using the affine calibration model we transform the uncalibrated score $s_j(\vec{q})$ of query $\vec{q}$ with a linear function 
    %
      \begin{equation}
       f_j(\vec{q}) = \alpha_j s_j(\vec{q}) + \beta_j,
       \label{eq:affine}
      \end{equation}
    %
    where 
    $ f_j(\vec{q})$ is the output calibrated score, and
    $\alpha_j$ and $\beta_j$ are scalar calibration parameters specific to each classifier $j$.
    In this work we use linear classifiers, hence substituting for $s_j(\vec{q})$ %in~\eqref{eq:affine} 
    the linear classifier from \eqref{eq:linear} results also in a linear calibrated classifier
    %
      \begin{align}
      \label{eq:linear2}
  %      \nonumber
  %        f_j(\vec{q}) = (c_j\vec{w}_j)^T \vec{q} + \beta_j
          f_j(\vec{q}) = \vec{\widetilde{w}}_j^T \vec{q} + \widetilde{b}_j,
          %\label{eq:affiine3}
       \end{align}
     %
  where $\vec{\widetilde{w}}_j = \alpha_j\vec{w}_j$ and $\widetilde{b}_j=\alpha_j b_j+\beta_j$. 
  Note that the calibrated classifier~\eqref{eq:linear2} has the same form as the original classifier~\eqref{eq:linear} and hence this representation does not require any additional storage compared to storing the original classifier. The question remains how to set the parameters $\alpha_j$ and $\beta_j$ of the calibration function~\eqref{eq:affine}, which is discussed next.

@@

      For the support vector machine classifier (SVM) training we use {\tt libsvm} \cite{libsvm}. We use the same $C_1$ and $C_2$ parameters for all per-exemplar classifiers, but find the optimal value of the parameters for each image representation by a cross-validation evaluating performance on a held out query set.

@@

      For the p-value calibration, we take the learnt classifier for each database image $j$ and compute its SVM score for all other database images to construct its empirical cumulative density function~\eqref{eq:empiricalCDF}. We keep only the top 1,000 values that, in turn, represent the calibration function. At query time, given the query descriptor $\vec{q}$, we compute the SVM score~\eqref{eq:linear} for each database image $j$, and compute its calibrated SVM score $f_j$~\eqref{eq:empiricalCDF}.

>>>>

\author{Petr Gron{\'a}t         \and
        Josef {S}ivic          \and
        Guillaume Obozinski   \and
        Tom{\'a}{\v s} Pajdla   %etc.
}

@@

\institute{Petr Gron{\'a}t \at
              2 rue Simone Iff \\
              75012 Paris, France\\
              \email{petr.gronat@inria.fr}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           %\and
}
\vspace{-5mm}
\date{Received: date / Accepted: date}

@@

  In this work, we also take advantage of geotags as an available form of supervision and investigate whether the place recognition problem can be cast as a classification task.
  {
    Learning visual classifiers has been investigated for landmark recognition~\cite{Li09} where consumer photographs were clustered into landmark classes based on geo-tags. 
    In this work we wish to recognize individual street locations rather than a small number of landmarks, and as a consequence have only a few (1-5) photographs
    %available for each place.   
    \textcolor{petr}{capturing the same location}.
    In particular, we train a classifier {\em for each location on the map} in a similar manner to per-exemplar classification in object recognition~\cite{Malisiewicz11}. 
    \textcolor{petr}{
      (By \textit{location} we mean a scene captured by a camera, hence using our terminology, a camera heading to the north captures different location than a camera heading to the south, even though these two cameras share the same GPS location. In the following text we use terms \textit{place} and \textit{location} interchangeably.)
    }
  }

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Per-location classifiers for place recognition}
\label{sec:classifiers}
%%%%%%%%%%%%%%%%%%%%%%%%
   We are given an image descriptor $\vec{x}_j$, one for each database image $j$. This representation can be a sparse tf-idf weighted bag-of-visual-words vector~\cite{Sivic03} or a dense compact descriptor such as the Fisher vector (FV)~\cite{Jegou12}. The goal is to learn 
   a score \textcolor{petr}{function}
   $f_j$ for each database image $j$, so that, at test time, given the descriptor $\vec{q}$ of the query image, we can either retrieve the correct target image as the image $j^*$ with the highest score
   % 
   \begin{equation}
   \label{eq:class}
    j^*=\operatorname*{arg\;max}_{j} f_j(\vec{q}) 
   \end{equation}
    %
   \noindent
   or use these scores to rank candidate images and use geometric verification to identify the correct location in an $n$-best list.
   Instead of approaching the problem directly as a large multiclass classification problem, we tackle the problem by learning a per-exemplar linear SVM classifier~\cite{Malisiewicz11}  for each database image $j$.
   Similar to~\cite{Knopp2010}, we use the available geotags to construct the negative set $\mathcal N_j$ for each image $j$. The negative set is constructed so as to concentrate difficult negative examples, i.e.\ from images that are far away from the location of image $j$ and similar to the target image as measured by the dot product between their feature vectors. The details of the construction procedure will be given in section~\ref{sec:experiments}.  The positive set $\mathcal P_j$ is represented by a single positive example
   , which is $\vec{x}_j$ itself. 
   \textcolor{petr}{
    For some locations the positive set can be expanded, details of this procedure will be given in~\ref{sub:paramsOfPerLoc}.
   }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%  
\section{Non-parametric calibration of the  SVM-scores from negative examples only}
\label{sec:calibration}
%%%%%%%%%%%%%%%%%%%%%%%%%%  
  In this section we describe a classifier calibration method that exploits the availability of large amounts of negative data, i.e.\ images from other far away locations in the database.
  In particular, the method estimates the significance of the \textcolor{petr}{classification} score of a test example compared to the typical \textcolor{petr}{classification} score of  the (plentifully available) negative examples. Intuitively, we will use a large dataset of negative examples to calibrate the individual classifiers so that they {\em reject the same number of negative examples} at each level of the calibrated score.   We will expand this idea in detail using the concepts from hypothesis testing.

   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \subsection{Calibration \emph{via} significance levels}
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      In the following, we view  the problem of deciding whether a query image matches a given location based on the corresponding SVM \textcolor{petr}{classification} score as a hypothesis testing problem. In particular, we appeal to ideas from the traditional frequentist hypothesis testing framework also known as Neyman-Pearson (NP) framework (see e.g.~\cite{casella2001statistical}, chap.~8).

@@

      In the NP framework, the \emph{significance level} of a score is measured by the p-value or equivalently by the value of the cumulative density function  (cdf) of the distribution of the negatives at a given score value. The cdf is the function $F_0$ defined by $F_0(s)=\PP(S_0\leq s)$, where $S_0$ is a random variable corresponding to the \textcolor{petr}{classification} scores of negative data (see figure~\ref{fig:qntExample} for an illustration of the relation between the cdf and the density of the function). The cdf 

@@

      where $(s_n)_{1\leq n \leq \Ncal }$ are the SVM \textcolor{petr}{classification} scores associated with $\Ncal$ negative examples used for calibration. Note that no positive examples are involved in the construction of the cumulative density function.

@@

      We also interpolate the empirical cdf between consecutive datapoints so that instead of being a staircase function it is a continuous piecewise linear function such as illustrated in figure~\ref{fig:calib}. Given a query, we first compute its SVM \textcolor{petr}{classification} score $s_q$ and then compute the calibrated probability, \textcolor{petr}{the score function} $f(q)=\hat{F}_0(s_q)$.
      We obtain a similar calibrated \textcolor{petr}{score function} $f_j(q)$ for each of the SVMs associated with each of the target locations, which can now be ranked.

@@

    For each trained place-specific classifier $s_j$ we construct the empirical cumulative density function~\eqref{eq:empiricalCDF} of \textcolor{petr}{classification} scores of the negative examples and keep only its top $K$ values. This can be done offline and the procedure is summarized in Algorithm \ref{alg:offline}. 

@@

   %%%%%%%%%%%%%%%%%%%%%%%
   \subsection{Discussion}
   \label{sec:discussion}
   %%%%%%%%%%%%%%%%%%%%%%% 
      It should be noted that basing the calibration only on the negative data has the advantage that we privilege precision over recall, which is justified given the imbalance of the available training data (many more negatives than positives).
      Indeed, since we are learning with a single positive example 
      \textcolor{petr}{(or a very few)}, 
      intuitively, we cannot guarantee that the learned partition of the space will generalize well to other positives, whose scores in the test set can potentially drop significantly. % (this is indeed what we observe in practice). 
      By contrast, since we are learning from a comparatively large number of negative examples, we can trust the fact that new negative examples will stay in the half-space containing the negative training set, so that their scores are very unlikely to be large. Our method is therefore based on the fact that we can measure reliably how surprising a high 
      \textcolor{petr}{classification}
      score would be if it was the score of a negative example. This exactly means that we can control false positives (type I error) reasonably well but not false negatives (type II error or equivalently the power of our test/classifier), exactly as in the Neyman-Pearson framework.

      An additional reason for not relying on positive examples for the calibration in our case is that (even if we had sufficiently many of them) the positive examples that we collect using location and geometric verification from the geotagged database
      \textcolor{petr}{(details in Section \ref{sub:paramsOfPerLoc})}
      typically have illumination conditions that are extremely similar to each other and not representative of the distribution of test positives which can have very different illuminations. This is because of the controlled nature of the capturing process of geotagged street-level imagery (e.g.\ Google Street View) used for experiments in this work. Close-by images are typically captured at a similar time (e.g. on the same day) and under similar imaging conditions.

@@

  \subsection{Affine calibration model}
    Using the affine calibration model we transform the uncalibrated \textcolor{petr}{classification}score $s_j(\vec{q})$ of query $\vec{q}$ with a linear function 
    %
      \begin{equation}
       f_j(\vec{q}) = \alpha_j s_j(\vec{q}) + \beta_j,
       \label{eq:affine}
      \end{equation}
    %
    where 
    $ f_j(\vec{q})$ is the output calibrated score, and
    $\alpha_j$ and $\beta_j$ are scalar calibration parameters specific to each classifier $j$.
    In this work we use linear classifiers, hence substituting for $s_j(\vec{q})$ %in~\eqref{eq:affine} 
    the linear classifier from \eqref{eq:linear} results also in a linear calibrated classifier
    %
      \begin{align}
      \label{eq:linear2}
  %      \nonumber
  %        f_j(\vec{q}) = (c_j\vec{w}_j)^T \vec{q} + \beta_j
          f_j(\vec{q}) = \vec{\widetilde{w}}_j^T \vec{q} + \widetilde{b}_j,
          %\label{eq:affiine3}
       \end{align}
     %
  where $\vec{\widetilde{w}}_j = \alpha_j\vec{w}_j$ and $\widetilde{b}_j=\alpha_j b_j+\beta_j$. 
  Note that the calibrated classifier~\eqref{eq:linear2} has the same form as the original classifier~\eqref{eq:linear} and hence this representation does not require any additional storage compared to storing the original classifier. The question remains how to set the parameters $\alpha_j$ and $\beta_j$ of the calibration function~\eqref{eq:affine}, which is discussed next.

@@

      For the support vector machine classifier (SVM) training we use {\tt libsvm} \cite{libsvm}. We use the same global $C_1$ and $C_2$ parameters for all per-exemplar classifiers, but find the optimal value of the parameters for each image representation by a cross-validation evaluating performance on a held out query set.

@@

      For the p-value calibration, we take the learnt classifier for each database image $j$ and compute its SVM \textcolor{petr}{classification} score for all other database images to construct its empirical cumulative density function~\eqref{eq:empiricalCDF}. We keep only the top 1,000 values that, in turn, represent the calibration function. At query time, given the query descriptor $\vec{q}$, we compute the SVM score~\eqref{eq:linear} for each database image $j$, and compute its calibrated SVM score $f_j$~\eqref{eq:empiricalCDF}.
